Supplementary Material: Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-11-12 si.tex
minor modifications to si and contour fig
made dots overlay dotted axes
minor issues with language in si adjusted
added explicit reference to section 4.5 in Morse and Feshbach

\subsection{General Solution to $\langle \tilde{n} | q^k | m \rangle$}
The matrix elements of \autoref{key} can be computed numerically, but analytic forms of the integral also exist. 
Due to the polynomial nature of $H_v(q)$, the product $H_n(q-\Delta) q^k H_m(q)$ will also be a polynomial, so that the integral in \autoref{key} will be of the form
\begin{eqnarray}\label{eq:Pnmkdef}
	\int_{-\infty}^\infty \mathrm{d}q \ P_{nmk}(q) \exp(-q^2 + q\Delta), \\
	P_{nmk}(q) \equiv \sum_\lambda^{n+m+k} a_\lambda q^\lambda
\end{eqnarray}
for some set coefficients $a_\lambda$ to be specified later. 
As a result, we can recast \autoref{key} into the linear combination of integrals
\begin{equation}\label{eq:analytic_form}
	\mel{\tilde{n}}{q^k}{m} = N_n N_m \exp(-\frac{\Delta^2}{2}) \sum_\lambda a_\lambda F_\lambda(\Delta),
\end{equation}
where
\begin{equation}\label{integral}
	F_w(\Delta) \equiv \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta).
\end{equation}
\autoref{integral} can be readily evaluated, as we now show. 
 When $w=0$, the integral is evaluated by completing the square:\footnote{$\int_{-\infty}^{\infty} \mathrm{d}x \exp(-(x-a)^2) = \sqrt{\pi}$}
\begin{equation}
	\begin{split}
		F_0(\Delta) & =  \int_{-\infty}^\infty \mathrm{d}q \exp\left(-q^2 + q\Delta\right)\\
		&= \int_{-\infty}^\infty \mathrm{d}q \exp(-(q^2 - q\Delta +\frac{\Delta^2}{4} - \frac{\Delta^2}{4}))\\
		&= \exp(\frac{\Delta^2}{4}) \int_{-\infty}^\infty \mathrm{d}q \exp(-(q - \frac{\Delta}{2})^2) \\
		&= \sqrt{\pi} \exp(\frac{\Delta^2}{4}).
	\end{split}
\end{equation}
For the general solution, \autoref{integral} is evaluated via repeated use of a well-known Gaussian identity
\begin{equation}\label{identity}
	\begin{split}
		F_w(\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta) = \partial_{\Delta} \int_{-\infty}^\infty \mathrm{d}q \ q^{w-1} \exp(-q^2 + q\Delta) \\
		&= \frac{\mathrm{d}}{\mathrm{d}\Delta} F_{w-1}(\Delta)
	\end{split}
\end{equation}
so we can write\footnote{
	Rearranging the Hermite polynomial generator so that $e^{-x^2} = (-1)^n H_n(x) \frac{\mathrm{d}^n}{\mathrm{d}x^n} e^{-x^2}$, setting $x \ = \ i \Delta /2$ and using the chain rule ($\frac{\mathrm{d}}{\mathrm{d}x} =  \frac{\mathrm{d}}{\mathrm{d}\Delta}  \frac{\mathrm{d\Delta}}{\mathrm{d}x}$) provides the final result in \autoref{eq:F_solved}.
	}
\begin{equation}\label{eq:F_solved}
	\begin{split}
		F_w(\Delta) &= \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} F_0(\Delta) \\
		&= \sqrt{\pi} \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} \exp(\frac{\Delta^2}{4}) \\
		&= \sqrt{\pi} \left( -\frac{i}{2} \right)^w \exp(\frac{\Delta^2}{4}) H_w\left(i\frac{\Delta}{2}\right).
	\end{split}
\end{equation}
In the last line we have used the generator definition of Hermite polynomials:  $H_n(x) \equiv (-1)^n e^{x^2} \frac{\mathrm{d}^n}{\mathrm{d}x^n} e^{-x^2}$.\cite{MorseFeshbach}
\vspace{3pt}

@@

In addition to the linewidth and offset impacting the shape of the HDFG response, the sign of the molecular parameters $M^{ge}_0, \Lambda^{ge}_0, \frac{\partial M^{ge}}{\partial Q} , \frac{\partial \Lambda^{ge}}{\partial Q}$ will significantly impact the HDFG spectrum.
Changing the sign of these parameters can change the sign of $A, B$ and alter the interference between the terms.  
As a specific example (\autoref{fig:chgsign}), we demonstrate the resultant HDFG spectrum for \begin{eqnarray}
	M^{ge}(q) = -1 + 0.04 q,  \\
	\Lambda^{ge}(q) = 1 + 0.04 q,
\end{eqnarray}
i.e., $M^{ge}_0 < 0$, $\Lambda^{ge}_0, \frac{\partial M^{ge}}{\partial Q} , \frac{\partial \Lambda^{ge}}{\partial Q} > 0$, keeping $\Delta = 0.5$.
Compared to Figure 3 of the main text, the contributions from the $A, B_1$ and $B_2$ terms to $\Re(\gamma)$ and $\Im(\gamma)$ significantly change.
In particular, $\Re(\gamma)$ and $\Im(\gamma)$ are of similar value in the post-resonance region, eliminating the destructive interference seen in Figure 3 of the main text.

@@

These effects have been recently observed in 1D and 2D coherent resonant Raman experiments.\cite{Fumero2020, Batignani2022}
When the sign of $\Delta$, $M(q)$ and $\Lambda(q)$ match, interference in the output lineshape can be observed.
Changing the absolute sign of these parameters can manipulate the interference (\autoref{fig:chgsign2}).
HDFG therefore has the ability to discern the absolute sign of $\Delta$ and assign offsets in potential energy wells for some systems.

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/changedeltaM.png}
	\caption{HDFG spectra for the system described by \autoref{Hamiltonian} with $M^{ge}(q) = M^{ge}_0 + 0.04 q,\ \Lambda^{ge}(q) = 1 + 0.04 q$. 
	The respective $\Delta, M^{0}_{ge}$ values are listed above the panels.
	Black trace: $\abs{A+B}$, blue trace: $\abs{A}$, orange trace: $\abs{B}$.
	The spectra are scaled to $\max{\abs{A+B}}$. 
	} 
	\label{fig:chgsign2}
\end{figure}
\newpage
\section{Evaluating the Impact of Static Inhomogeneity on HDFG Spectra}
In this section, the integral
\begin{equation}\label{generalcontour}
	\begin{split}
		\gamma_{ijkl} &= \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi)\\
		&= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)} \\
		&= \frac{\eta_{ijkl} \sigma}{\pi} \int_{-\infty}^\infty \mathrm{d}\xi\frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} + c\xi}\\
	\end{split}
\end{equation}
is evaluated for $c=-1$ (electronic state and vibrational state energy fluctuations are anti-correlated) and $c=1$ (electronic state and vibrational state energy fluctuations are correlated).

Integrals such as \autoref{generalcontour} can be evaluated by using the Residue theorem.\cite{MorseFeshbach}
For clarity, we briefly outline the technique.
Consider the integral
\begin{equation}\label{eq:integral}
	\int_{-\infty}^{\infty} \mathrm{d}x f(x).
\end{equation}
Assume \autoref{eq:integral} converges.
If $f(z)$ is analytic on the upper half of the complex plane (except for poles on the real axis), where $z$ is a complex variable, and $f(z)$ asymptotically approaches zero with the form $\abs{Az}^{-m}$, $m > 1$, as $z \rightarrow \infty$, then
\begin{subequations}\label{eq:generalcontour}
	\begin{equation}
		\int_{-\infty}^{\infty}  \mathrm{d}x f(x) = 2\pi i \sum_k \Res[f(z_k)],
	\end{equation}
where $\Res[f(z_k)]$ refers to the residue of $f(z)$ in the upper half of the complex plane at the pole $z_k$ and the sum is taken over all the poles, i.e., non-removable singularities, of $f(z)$.\cite{MorseFeshbach}
Note that the residue of an n$^{\text{th}}$ order pole at $z = z_k$ is evaluated as
	\begin{equation}
		\Res[f(z_k)] = \lim_{z \rightarrow z_k} \frac{1}{\left(n-1\right)!} \frac{\mathrm{d}^{n-1}}{\mathrm{d}z^{n-1}} \left[f(z) (z-z_k)^n \right].
	\end{equation}
For first order poles (the cases considered below), $n=1$, so that 
	\begin{equation}
		\int_{-\infty}^{\infty} \mathrm{d}x f(x) = 2\pi i \sum_k \lim_{z \rightarrow z_k} f(z) (z-z_k).
	\end{equation}
\end{subequations}
Note that in \autoref{eq:generalcontour}, the integrals are evaluated along a contour which wraps all the poles in the positive half of the complex plane.

@@

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/contour.png}
	\caption{Poles (blue dots) and contours (black lines) used to evaluate \autoref{generalcontour} for (a) anti-correlated ($c=-1$) and (b) correlated ($c=1$) modes. 
	} 
	\label{fig:contours}
\end{figure}

>>>>

\subsection{General Solution to $\langle \tilde{n} | q^k | m \rangle$}
The matrix elements of \autoref{key} can be computed numerically, but analytic forms of the integral also exist. 
Due to the polynomial nature of $H_v(q)$, the product $H_n(q-\Delta) q^k H_m(q)$ will also be a polynomial, so that the integral in \autoref{key} will be of the form
\begin{eqnarray}\label{eq:Pnmkdef}
	\int_{-\infty}^\infty \mathrm{d}q \ P_{nmk}(q) \exp(-q^2 + q\Delta), \\
	P_{nmk}(q) \equiv \sum_\lambda^{n+m+k} a_\lambda q^\lambda
\end{eqnarray}
for some set coefficients $a_\lambda$ to be specified later. 
As a result, we can recast \autoref{key} into the linear combination of integrals
\begin{equation}\label{eq:analytic_form}
	\mel{\tilde{n}}{q^k}{m} = N_n N_m \exp(-\frac{\Delta^2}{2}) \sum_\lambda a_\lambda F_\lambda(\Delta),
\end{equation}
where
\begin{equation}\label{integral}
	F_w(\Delta) \equiv \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta).
\end{equation}
\autoref{integral} can be readily evaluated, as we now show. 
 When $w=0$, the integral is evaluated by completing the square:\footnote{$\int_{-\infty}^{\infty} \mathrm{d}x \exp(-(x-a)^2) = \sqrt{\pi}$}
\begin{equation}
	\begin{split}
		F_0(\Delta) & =  \int_{-\infty}^\infty \mathrm{d}q \exp\left(-q^2 + q\Delta\right)\\
		&= \int_{-\infty}^\infty \mathrm{d}q \exp(-(q^2 - q\Delta +\frac{\Delta^2}{4} - \frac{\Delta^2}{4}))\\
		&= \exp(\frac{\Delta^2}{4}) \int_{-\infty}^\infty \mathrm{d}q \exp(-(q - \frac{\Delta}{2})^2) \\
		&= \sqrt{\pi} \exp(\frac{\Delta^2}{4}).
	\end{split}
\end{equation}
For the general solution, \autoref{integral} is evaluated via repeated use of a well-known Gaussian identity
\begin{equation}\label{identity}
	\begin{split}
		F_w(\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta) = \partial_{\Delta} \int_{-\infty}^\infty \mathrm{d}q \ q^{w-1} \exp(-q^2 + q\Delta) \\
		&= \frac{\mathrm{d}}{\mathrm{d}\Delta} F_{w-1}(\Delta)
	\end{split}
\end{equation}
so we can write\footnote{
	Rearranging the Hermite polynomial generator so that $\frac{\mathrm{d}^n}{\mathrm{d}x^n} e^{-x^2} = (-1)^n H_n(x) e^{-x^2}$, setting $x \ = \ i \Delta /2$ and using the chain rule ($\frac{\mathrm{d}}{\mathrm{d}x} =  \frac{\mathrm{d}}{\mathrm{d}\Delta}  \frac{\mathrm{d\Delta}}{\mathrm{d}x}$) provides the final result in \autoref{eq:F_solved}.
	}
\begin{equation}\label{eq:F_solved}
	\begin{split}
		F_w(\Delta) &= \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} F_0(\Delta) \\
		&= \sqrt{\pi} \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} \exp(\frac{\Delta^2}{4}) \\
		&= \sqrt{\pi} \left( -\frac{i}{2} \right)^w \exp(\frac{\Delta^2}{4}) H_w\left(i\frac{\Delta}{2}\right).
	\end{split}
\end{equation}
In the last line we have used the generator definition of Hermite polynomials:  $H_n(x) \equiv (-1)^n e^{x^2} \frac{\mathrm{d}^n}{\mathrm{d}x^n} e^{-x^2}$.\cite{MorseFeshbach}
\vspace{3pt}

@@

In addition to the linewidth and offset impacting the shape of the HDFG response, the sign of the molecular parameters $M^{ge}_0, \Lambda^{ge}_0, \frac{\partial M^{ge}}{\partial Q} , \frac{\partial \Lambda^{ge}}{\partial Q}$ will significantly impact the HDFG spectrum.
Changing the sign of these parameters can change the sign of $A, B$ and alter the interference between the terms.  
As a specific example (\autoref{fig:chgsign}), we demonstrate the resultant HDFG spectrum for \begin{eqnarray}
	M^{ge}(q) = -1 + 0.04 q,  \\
	\Lambda^{ge}(q) = 0.1 + 0.004 q,
\end{eqnarray}
i.e., $M^{ge}_0 < 0$, $\Lambda^{ge}_0, \frac{\partial M^{ge}}{\partial Q} , \frac{\partial \Lambda^{ge}}{\partial Q} > 0$, keeping $\Delta = 0.5$.
Compared to Figure 3 of the main text, the contributions from the $A, B_1$ and $B_2$ terms to $\Re(\gamma)$ and $\Im(\gamma)$ significantly change.
In particular, $\Re(\gamma)$ and $\Im(\gamma)$ are of similar value in the post-resonance region, eliminating the destructive interference seen in Figure 3 of the main text.
Similar effects have been recently observed in 1D and 2D coherent resonant Raman experiments.\cite{Fumero2020, Batignani2022}
When the sign of $\Delta$, $M(q)$ and $\Lambda(q)$ match, interference in the output lineshape can be observed.
Changing the absolute sign of these parameters can manipulate the interference (\autoref{fig:chgsign2}).
HDFG therefore has the ability to discern the absolute sign of $\Delta$ and assign offsets in potential energy wells for some systems.

@@

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/changedeltaM.png}
	\caption{HDFG spectra for the system described by \autoref{Hamiltonian} with $M^{ge}(q) = M^{ge}_0 + 0.04 q,\ \Lambda^{ge}(q) = 0.01 + 0.004 q$. 
	The respective $\Delta, M^{0}_{ge}$ values are listed above the panels.
	Black trace: $\abs{A+B}$, blue trace: $\abs{A}$, orange trace: $\abs{B}$.
	The spectra are scaled to $\max{\abs{A+B}}$. 
	} 
	\label{fig:chgsign2}
\end{figure}
\newpage
\section{Evaluating the Impact of Static Inhomogeneity on HDFG Spectra}
In this section, the integral
\begin{equation}\label{generalcontour}
	\begin{split}
		\gamma_{ijkl} &= \int_{-\infty}^\infty \mathrm{d}\xi P(\xi) \gamma_{ijkl}(\xi)\\
		&= \frac{\sigma}{\pi}\int_{-\infty}^\infty \mathrm{d}\xi \frac{1}{\sigma^2 + \xi^2} \frac{\eta_{ijkl}}{\left(\Delta_{ga} - \xi\right)\left(\Delta_{ba}+ c\xi\right)} \\
		&= \frac{\eta_{ijkl} \sigma}{\pi} \int_{-\infty}^\infty \mathrm{d}\xi\frac{1}{(\xi + i\sigma)(\xi - i\sigma)} \frac{1}{\Delta_{ga} - \xi} \frac{1}{\Delta_{ba} + c\xi}\\
	\end{split}
\end{equation}
is evaluated for $c=-1$ (electronic state and vibrational state energy fluctuations are anti-correlated) and $c=1$ (electronic state and vibrational state energy fluctuations are correlated).

Integrals such as \autoref{generalcontour} can be evaluated by using the Residue theorem.\cite{MorseFeshbach}
For clarity, we briefly outline the technique.\footnote{See \S 4.5 of Morse and Feshbach \cite{MorseFeshbach} for more details.} 
Consider the integral
\begin{equation}\label{eq:integral}
	\int_{-\infty}^{\infty} \mathrm{d}x f(x).
\end{equation}
Assume \autoref{eq:integral} converges.
If $f(z)$ is analytic on the upper half of the complex plane (except for poles not on the real axis), where $z$ is a complex variable, and $\abs{f(z)}$ asymptotically approaches zero with the form $\abs{Az}^{-m}$, $m > 1$, as $\abs{z} \rightarrow \infty$, then
\begin{subequations}\label{eq:generalcontour}
	\begin{equation}
		\int_{-\infty}^{\infty}  \mathrm{d}x f(x) = 2\pi i \sum_k \Res[f(z_k)],
	\end{equation}
where $\Res[f(z_k)]$ refers to the residue of $f(z)$ at the pole $z_k$ and the sum is taken over all the poles of $f(z)$ in the upper half of the complex plane.\cite{MorseFeshbach}
Note that the residue of an n$^{\text{th}}$ order pole at $z = z_k$ is evaluated as
	\begin{equation}
		\Res[f(z_k)] = \lim_{z \rightarrow z_k} \frac{1}{\left(n-1\right)!} \frac{\mathrm{d}^{n-1}}{\mathrm{d}z^{n-1}} \left[f(z) (z-z_k)^n \right].
	\end{equation}
For first order poles (the cases considered below), $n=1$, so that 
	\begin{equation}
		\int_{-\infty}^{\infty} \mathrm{d}x f(x) = 2\pi i \sum_k \lim_{z \rightarrow z_k} f(z) (z-z_k).
	\end{equation}
\end{subequations}
Note that in \autoref{eq:generalcontour}, the integrals are evaluated along a counter-clockwise contour that wraps all the poles in the positive half of the complex plane.

@@

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/contour.png}
	\caption{Poles (blue dots) and contours (black lines) used to evaluate \autoref{generalcontour} for (a) anti-correlated ($c=-1$) and (b) correlated ($c=1$) modes. 
	The contours are oriented counter-clockwise.
	} 
	\label{fig:contours}
\end{figure}
